---
title: "Chapter 2 HW"
author: "Nehemya McCarter-Ribakoff"
date: "14 February 2017"
output: 
  html_document: 
    fig_height: 4
    highlight: pygments
    theme: spacelab
---

```{r setup, include=FALSE}
# DO NOT ALTER CODE IN THIS CHUNK
knitr::opts_chunk$set(echo = TRUE)
library(plyr)
library(ggplot2)
```

* * *

## Conceptual Questions


#### Exercise 2: Identify each as classification or regression, prediction
                 or inference. Provide n and p
    
    (a) We collect a set of data on the top 500 firms in the US. For each
    firm we record profit, number of employees, industry and the
    CEO salary. We are interested in understanding which factors
    affect CEO salary.
    
    Since we are looking for factors, or categories, this is classification.
    We are not building a model for prediction, only studying patterns, so
    this is inference.
    n = 500, p = {profit, number of employees, industry, and CEO salary}
    
    (b) We are considering launching a new product and wish to know
    whether it will be a success or a failure. We collect data on 20
    similar products that were previously launched. For each product
    we have recorded whether it was a success or failure, price
    charged for the product, marketing budget, competition price,
    and ten other variables.
    
    Since products are labeled either success or failure, this is classification.
    We are determining if our product will be a success, so this is prediction.
    n = 20, p = {success or failure, price of product, marketing budget,
                 competition budget, 10 other...}
                 
    (c) We are interested in predicting the % change in the US dollar in
    relation to the weekly changes in the world stock markets. Hence
    we collect weekly data for all of 2012. For each week we record
    the % change in the dollar, the % change in the US market,
    the % change in the British market, and the % change in the
    German market.
    
    We are looking for a percent change. This is regression.
    We are making a prediction model.
    n = 52, p = {Δdollar, ΔUS, ΔBritish, ΔGerman}


#### Exercise 4: You will now think of some real-life applications for statistical learning.
         
    (a) Describe three real-life applications in which classification might
    be useful. Describe the response, as well as the predictors. Is the
    goal of each application inference or prediction? Explain your
    answer.
    
    1. Mobile phone application to determine the species of a plant in a given
       photo based off a training dataset.
       
       predictors = {plant.shape, plant.color, plant.structure, ...}
       Prediction model
       The model looks for a photo or species in its database that matches the
       given photo as closely as possible. This is an image classifier.
       
    2. Virtual doctor providing diagnoses based on symptoms.
    
       predictors = {body.temp, inflammation(if any), reported.symptoms, ...}
       Goal: prediction
       The output of our model is an element from the set of all diseases.
       
    3. Determine if decreasing size of fish population is correlated with
       region, species, diet, etc.
       
       predictors = {region, species, diet, etc.}
       Inference model.
       Our outcomes are discrete -- either there is a correlation or there isn't.
       Additionally, we are only checking if there is a pattern. No output is
       generated, so this is inference.
     
    (b) Describe three real-life applications in which regression might
    be useful. Describe the response, as well as the predictors. Is the
    goal of each application inference or prediction? Explain your
    answer.
    
    1. Find the probability a tornado will occur based on climate data.
       predictors = {atmospheric.pressure, temperature, wind.data, ...}
       Goal: prediction
       Because a probability is a continuous value, this is a regression problem
    
    2. Estimate the property value of a given plot of land
       predictors = {location, size, structures, ...}
       Goal: prediction
       A property value is continuous, but because the price at which a plot will
       sell is, to a degree, arbitrary, this problem is prediction, not
       inference.
    
    3. Determine how much each factor impacts on property sales.
       predictors = {location, size, structures, ...}
       Goal: inference
       Our scenario is similar to the last example, but we are only looking
       for a better understanding of the relationship between our predictors
       and outcome. Since our end goal is quantitative ("how much"), this is
       also an inference problem.

* * *

## Applied Questions

#### Exercise 8: 
(a) add code to the chuck.



```{r}

```

(b) make comments if needed.

#### Exercise 10: 




* * *

## Teamwork report

<!--
For each team member, write their name, whether they attended the lab session, 
whether they were the author for this week, and their contribution to the lab 
report (as a percentage). The percentages should add up to 100.
-->

| Team member             | Conceptual | Applied  | Contribution % |
|:------------------------|:----------:|:--------:|---------------:|
| Nehemya                 |    Yes     |   Yes    | 100%           |
| Total                   |            |          | 100%           |