---
title: "Chapter 9 HW"
author: "Nehemya McCarter-Ribakoff"
date: "11 May 2017"
output: 
  html_document: 
    fig_height: 4
    highlight: pygments
    theme: spacelab
---

```{r setup, include=FALSE}
# DO NOT ALTER CODE IN THIS CHUNK
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(ggplot2)
```

* * *

## Conceptual Questions


#### Exercise 1: This problem involves hyperplanes in two dimensions.

(a) Sketch the hyperplane 1 + 3X<sub>1</sub> − X<sub>2</sub> = 0. Indicate the set of points for which 1 + 3X<sub>1</sub> − X<sub>2</sub> > 0, as well as the set of points for which 1 + 3X<sub>1</sub> − X<sub>2</sub> < 0.

(b) On the same plot, sketch the hyperplane −2 + X1 + 2X2 = 0. Indicate the set of points for which −2 + X<sub>1</sub> + 2X<sub>2</sub> > 0, as well as the set of points for which −2 + X<sub>1</sub> + 2X<sub>2</sub> < 0.

#### Exercise 2: We have seen that in p = 2 dimensions, a linear decision boundary takes the form β<sub>0</sub>+β<sub>1</sub>X<sub>1</sub>+β<sub>2</sub>X<sub>2</sub> = 0. We now investigate a non-linear decision boundary.

(a) Sketch the curve

(1 + X<sub>1</sub>)<sup>2</sup> + (2 − X<sub>2</sub>)<sup>2</sup> = 4.

(b) On your sketch, indicate the set of points for which

(1 + X<sub>1</sub>)<sup>2</sup> + (2 − X<sub>2</sub>)<sup>2</sup> > 4,

as well as the set of points for which

(1 + X<sub>1</sub>)<sup>2</sup> + (2 − X<sub>2</sub>)<sup>2</sup> ≤ 4.

(c) Suppose that a classifier assigns an observation to the blue class if

(1 + X<sub>1</sub>)<sup>2</sup> + (2 − X<sub>2</sub>)<sup>2</sup> > 4,

and to the red class otherwise. To what class is the observation (0, 0) classified? (−1, 1)? (2, 2)? (3, 8)?

(d) Argue that while the decision boundary in (c) is not linear in terms of X<sub>1</sub> and X<sub>2</sub>, it is linear in terms of X<sub>1</sub>, X<sub>1</sub><sup>2</sup> , X<sub>2</sub>, and X<sup>2</sup><sub>2</sub>.

* * *

## Applied Questions

#### Exercise 7: In this problem, you will use support vector approaches in order to
predict whether a given car gets high or low gas mileage based on the
Auto data set.
(a) Create a binary variable that takes on a 1 for cars with gas mileage above the median, and a 0 for cars with gas mileage below the median.

(b) Fit a support vector classifier to the data with various values of cost, in order to predict whether a car gets high or low gas mileage. Report the cross-validation errors associated with different values of this parameter. Comment on your results.

(c) Now repeat (b), this time using SVMs with radial and polynomial basis kernels, with different values of gamma and degree and cost. Comment on your results.

(d) Make some plots to back up your assertions in (b) and (c).

Hint: In the lab, we used the plot() function for svm objects only in cases with p = 2. When p > 2, you can use the plot() function to create plots displaying pairs of variables at a time. Essentially, instead of typing

> plot(svmfit , dat)
where svmfit contains your fitted model and dat is a data frame
containing your data, you can type
> plot(svmfit , dat , x1∼x4)
in order to plot just the first and fourth variables. However, you
must replace x1 and x4 with the correct variable names. To find
out more, type ?plot.svm.

```{r}

```


* * *

## Teamwork report

| Team member             | Conceptual | Applied  | Contribution % |
|:------------------------|:----------:|:--------:|---------------:|
| Nehemya                 |    Yes     |   Yes    | 100%           |
| Total                   |            |          | 100%           |